---
title: "Cost Conversion Resource"
author: "Silvia P. Canelón"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE}
# loading packages----
pkgs = c("tidyverse", "janitor", "gt", "DT", "here", "glue", "fontawesome")
pacman::p_load(char = pkgs, install = FALSE, update = FALSE)

# setting options----
knitr::opts_chunk$set(echo = TRUE, fig.retina = 3, 
                      warning=FALSE, message=FALSE)
options(scipen=999)
```
# `r fa("arrow-circle-down")` Jump down to the [Cost conversion chart](#cost-conversion-chart)


# Data import

```{r}
# setting data directory
dirData = here::here("data-raw/")

# World Bank Country and Lending Groups
classification_raw <- 
  readxl::read_excel(
    path = glue(dirData, "WorldBank-Country-Income-Class.xls"),
    sheet = "List of economies",
    range = "A5:I224") %>%
  janitor::clean_names()
glimpse(classification_raw)

# Purchasing Power Parity
gdp_by_ppp_raw <- 
 read_csv(glue(dirData, "ICP-2017-GDP-PPP-Data.csv"),
          col_types = "fffffcn") %>%
  janitor::clean_names()
glimpse(gdp_by_ppp_raw)

gdp_by_ppp_meta_raw <- 
  read_csv(glue(dirData, "ICP-2017-GDP-PPP-Metadata.csv")) %>%
  janitor::clean_names()
```

# Data wranging

```{r}
classification <- 
  classification_raw %>%
  filter(economy != "x") %>%
  select(-x_1) %>%
  remove_empty(c("rows", "cols")) %>%
  rename("economy_name"="economy", "economy_code"="code") %>%
  mutate_at(vars(region, income_group, lending_category), factor) %>%
  mutate(income_group = factor(income_group, levels = unique(income_group)),
         income_group = fct_relevel(income_group, 
                                    "Low income", "Lower middle income"))
glimpse(classification)

gdp_by_ppp <-
  gdp_by_ppp_raw %>%
    select(-series_code) %>%
    filter(!is.na(country_code)) %>%
    rename("economy_name"="country_name", "economy_code"="country_code",
           "value"="x2017_yr2017") %>%
    mutate(value = round(value, 2),
           series_name = 
             ifelse(series_name == "1000000:GROSS DOMESTIC PRODUCT", 
                    "GDP", NA)) 
glimpse(gdp_by_ppp)
```

# Data trimming

```{r}
classification_trim <-
  classification %>%
  select(economy_name, economy_code, region, income_group)
glimpse(classification_trim)

gdp_by_ppp_trim <-
  gdp_by_ppp %>%
    select(economy_name, economy_code, 
           classification_code, value) %>%
  pivot_wider(names_from = classification_code, values_from = value) %>%
  rename("PPP_GDPCap"="PCAP.PP", "PPP"="PPPGlob")
glimpse(gdp_by_ppp_trim)
```

# Joining classification data with PPP data

```{r}
gdp_ppp_full <- classification_trim %>% left_join(gdp_by_ppp_trim)
glimpse(gdp_ppp_full)
```
# Data viz

```{r}
library(viridis)

# boxplot ----
gdp_ppp_full %>%
  ggplot(aes(x = PPP_GDPCap, y = income_group)) +
  geom_boxplot(aes(fill = income_group)) +
  scale_fill_viridis_d() +
  scale_x_continuous(breaks=seq(0, 130000, by = 20000)) +
  theme_minimal() +
  theme(text = element_text(family = "Source Sans Pro"),
        plot.title.position = "plot",
        legend.position = "none",
        panel.grid.minor.x = element_blank()) +
  labs(x = "PPP-based GDP per capita (US$)", y = NULL,
       title = "World economies grouped into income groups",
       subtitle = "Utilizing Gross Domestic Product (GDP, per capita) adjusted for Purchasing Power Parity (PPP)",
       caption = "Data source: ICP 2017 -- Last updated: Oct. 01, 2020")

# scatterplot ----
gdp_ppp_full %>%
  ggplot(aes(x = PPP_GDPCap, y = log10(PPP))) +
  geom_point(aes(color = income_group)) +
  scale_colour_viridis_d(name = "Income group") +
  scale_x_continuous(breaks=seq(0, 130000, by = 20000)) +
  theme_minimal() +
  theme(text = element_text(family = "Source Sans Pro"),
        plot.title.position = "plot",
        legend.position = c(1,1),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.background = element_rect(color = "white"),
        panel.grid.minor.x = element_blank()) +
  labs(x = "PPP-based GDP per capita (US$)",
       title = "World economies grouped into income groups",
       subtitle = "Utilizing Gross Domestic Product (GDP, per capita) adjusted for Purchasing Power Parity (PPP)",
       caption = "Data source: ICP 2017 -- Last updated: Oct. 01, 2020")
```

# Cost conversion

## Example

```{r}
# summarizing GDP (PPP-adjusted) per income_group
gdp_ppp_summary <-
gdp_ppp_full %>%
  na.omit(PPP_GDPCap) %>%
  group_by(income_group) %>%
  summarize(avg_GDP = mean(PPP_GDPCap), sd_GDP = sd(PPP_GDPCap)) %>%
  mutate(norm_avg_GDP = avg_GDP/max(avg_GDP)) %>%
  arrange(desc(avg_GDP)) %>%
  mutate_if(is.numeric, round, 2)
gdp_ppp_summary %>% gt()

# adding sponsorship level example
cost_conversion <-
  gdp_ppp_summary %>%
  select(-(avg_GDP:sd_GDP)) %>%
  rename("conversion_factor"="norm_avg_GDP") %>%
  mutate(platinum = 15000,
         gold = 10000,
         silver = 7500,
         bronze = 3000,
         network = 1000) %>%
  mutate(platinum = platinum*conversion_factor, # adjusts by the conversion factor
         gold = gold*conversion_factor,
         silver = silver*conversion_factor,
         bronze = bronze*conversion_factor,
         network = network*conversion_factor)
```

**Modifications:**

- To create sponsorship packages, `income_group`s "Low income" and "Lower middle income" were combined into one category: "Lower income"

```{r}
# dropping the "Low income" category and reclassifying "Lower middle income"
# to create sponsorship packages
cost_conversion_trim <-
  cost_conversion %>%
  filter(income_group != "Low income") %>%
  mutate(income_group = fct_recode(income_group,
                                   "Lower income" = "Lower middle income"))
cost_conversion_trim %>% 
  gt() %>%
  tab_header(
    title = md("Cost Conversions by Global Income Group"),
    subtitle = md("Adjustments by GDP and PPP (US$)")
  ) %>%
  cols_label(
    income_group = "Global Income Group",
    conversion_factor = "Conversion Factor",
    platinum = "Platinum",
    gold = "Gold",
    silver = "Silver",
    bronze = "Bronze",
    network = "Network"
  ) %>%
  fmt_currency(
    columns = c("platinum", "gold", "silver", "bronze", "network"),
    currency = "USD",
    decimals = 0
  )
```

## Cost conversion chart

Prices listed are in USD

```{r}
# recoding "Low income" and "Lower middle income" as "Lower income"
# to create sponsorship packages

classification_trim %>% 
  mutate(income_group = fct_recode(income_group,
                                   "Lower income" = "Low income",
                                   "Lower income" = "Lower middle income")) %>%
  left_join(cost_conversion_trim) %>%
  mutate_at(vars(conversion_factor:network), factor) %>%
  select(-conversion_factor, -economy_code) %>%
  DT::datatable(style = "bootstrap", 
                filter = 'top', 
                options = list(
                  pageLength = 15,
                  autoWidth = TRUE),
                colnames = c('Country/Economy', 'Region', 'Global Income Group',
                             'Platinum', 'Gold', 'Silver', 'Bronze', 'Network')) %>%
  DT::formatCurrency(c("platinum", "gold", "silver", "bronze", "network"),
                     digits = 0)
```

